---
title: "Group Project"
author: "Maggie Carttar, Yue Kuang, Emma Livingston"
date: "November 6, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(readr)
library(tidyverse)
```

## Importing data
```{r}
ma_schools<-read.csv("ma_schools.csv")
```


```{r}
pruned_ma_schools <- ma_schools %>%
  select(X10_Enrollment, TOTAL_Enrollment, First.Language.Not.English, X..English.Language.Learner, English.Language.Learner, X..English.Language.Learner, Students.With.Disabilities, X..Students.With.Disabilities, High.Needs, X..High.Needs, Economically.Disadvantaged, X..Economically.Disadvantaged, X..African.American, X..Asian, X..Hispanic, X..White, X..Native.American, X..Native.Hawaiian..Pacific.Islander, X..Multi.Race..Non.Hispanic, X..Males, X..Females, Total...of.Classes, Average.Class.Size, Number.of.Students, Salary.Totals, Average.Salary, FTE.Count, In.District.Expenditures, Total.In.district.FTEs, Average.In.District.Expenditures.per.Pupil, Total.Expenditures, Total.Pupil.FTEs, Average.Expenditures.per.Pupil, X..Graduated, X..Still.in.School, X..Non.Grad.Completers, X..GED, X..Dropped.Out, X..Permanently.Excluded, X..Attending.College, MCAS_10thGrade_English_A.., MCAS_10thGrade_English_CPI, MCAS_10thGrade_Math_P.., MCAS_10thGrade_English_Incl..in.SGP..., MCAS_10thGrade_English_NI.., MCAS_10thGrade_English_P.., MCAS_10thGrade_English_P.A.., MCAS_10thGrade_English_SGP, MCAS_10thGrade_English_Stud..Incl..., MCAS_10thGrade_English_W.F.., MCAS_10thGrade_Math_A.., MCAS_10thGrade_Math_CPI, MCAS_10thGrade_Math_Incl..in.SGP..., MCAS_10thGrade_Math_NI.., MCAS_10thGrade_Math_P.., MCAS_10thGrade_Math_P.A.., MCAS_10thGrade_Math_SGP, MCAS_10thGrade_Math_Stud..Incl..., MCAS_10thGrade_Math_W.F.., X..MCAS_10thGrade_English_A, X..MCAS_10thGrade_English_NI, X..MCAS_10thGrade_English_P, X..MCAS_10thGrade_English_P.A, X..MCAS_10thGrade_English_W.F, X..MCAS_10thGrade_Math_A, X..MCAS_10thGrade_Math_NI, X..MCAS_10thGrade_Math_P, X..MCAS_10thGrade_Math_P.A, X..MCAS_10thGrade_Math_W.F)
```

There are 1861 schools (observations) and 67 variables (reduced from 302). All MCAS information is based on 10th graders. The variables are as follows:
Maggie: 1-22
Yue: 23 - 45
Emma: 46 - 67
40. `MCAS_10thGrade_English_A..` : an `int` of Advanced scores in English.  
41. `MCAS_10thGrade_English_CPI` : the `num` measure of Composite Performance Index for English.
42. `MCAS_10thGrade_Math_P..` : an `int` of Proficient scores in Math.
43. `MCAS_10thGrade_English_Incl..in.SGP...` : an `int` of the students included in the Student Growth Percentile in English.
44. `MCAS_10thGrade_English_NI..` : an `int` of Needs Improvement scores in English.
45. `MCAS_10thGrade_English_P..` : an `int` of Proficient scores in English.
46. `MCAS_10thGrade_English_P.A..` : an `int` of Proficient or Advanced scores in English.
47. `MCAS_10thGrade_English_SGP` : a `num` measure of Student Growth Percentile in English.
48. `MCAS_10thGrade_English_Stud..Incl...` : an `int` of the number of students included in the measure of SGP.
49. `MCAS_10thGrade_English_W.F..` : an `int` of Warning/Failing scores in English.
50. `MCAS_10thGrade_Math_A..` : an `int` of Advanced scores in Math.
51. `MCAS_10thGrade_Math_CPI` : a `num` measure of Composite Performance Index for Math.
52. `MCAS_10thGrade_Math_Incl..in.SGP...` : an `int` of the students included in the Student Growth Percentile in Math.
53. `MCAS_10thGrade_Math_NI..` : an `int` of Needs Improvement scores in Math.
54. `MCAS_10thGrade_Math_P.A..` : an `int` of Proficient or Advanced scores in Math.
55. `MCAS_10thGrade_Math_SGP`: a `num` measure of Student Growth Percentile in English.
56. `MCAS_10thGrade_Math_Stud..Incl...` : an `int` of the number of students included in the measure of SGP.
57. `MCAS_10thGrade_Math_W.F..` : an `int` of Warning/Failing scores in Math.
58. `X..MCAS_10thGrade_English_A` : an `int` percent of Advanced scores in English.
59. ` X..MCAS_10thGrade_English_NI` : an `int` percent of Needs Improvement scores in English.
60. `X..MCAS_10thGrade_English_P` : an `int` percent of Proficient scores in English.
61. `X..MCAS_10thGrade_English_P.A` : an `int` percent of Proficient or Advanced scores in English.
62. `X..MCAS_10thGrade_English_W.F` : an `int` percent of Warning/Failing scores in English.
63. `X..MCAS_10thGrade_Math_A` : an `int` percent of Advanced scores in Math.
64. `X..MCAS_10thGrade_Math_NI` : an `int` percent of Needs Improvement scores in Math.
65. `X..MCAS_10thGrade_Math_P` : an `int` percent of Proficient scores in Math.
66. `X..MCAS_10thGrade_Math_P.A` : an `int` percent of Proficient or Advanced scores in Math. 
67. `X..MCAS_10thGrade_Math_W.F` : an `int` percent of Warning/Failing scores in Math.





##get rid of schools without tenth grade 
```{r}
subset(pruned_ma_schools,
       !is.na(pruned_ma_schools$MCAS_10thGrade_Math_A..))
```

```{r}
str(pruned_ma_schools, give.attr = FALSE)
```

#Key Next Steps
1. Eliminate the variables that we know off-hand are not useful in our analysis. 
2. Test for collinearity between variables that contain related information (for example, we will likely not include both Salary Totals and Average Salary, which both pertain to the salary of teachers).
3. Create a "Percent Proficient" variable to be used as our response variable out of the categories provided. 
4. Rename some of the variables so they make more sense. We should also consider re-ordering them.
