---
title: "Group Project"
author: "Maggie Carttar, Yue Kuang, Emma Livingston"
date: "November 6, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(readr)
library(tidyverse)
```

## Importing data
```{r}
ma_schools<-read.csv("ma_schools.csv")
```

##Look at format of each variable
```{r}
str(ma_schools_pruned, give.attr = FALSE)
```


```{r}
pruned_ma_schools <- ma_schools %>%
  select(X10_Enrollment, TOTAL_Enrollment, First.Language.Not.English, X..English.Language.Learner, English.Language.Learner, X..English.Language.Learner, Students.With.Disabilities, X..Students.With.Disabilities, High.Needs, X..High.Needs, Economically.Disadvantaged, X..Economically.Disadvantaged, X..African.American, X..Asian, X..Hispanic, X..White, X..Native.American, X..Native.Hawaiian..Pacific.Islander, X..Multi.Race..Non.Hispanic, X..Males, X..Females, Total...of.Classes, Average.Class.Size, Number.of.Students, Salary.Totals, Average.Salary, FTE.Count, In.District.Expenditures, Total.In.district.FTEs, Average.In.District.Expenditures.per.Pupil, Total.Expenditures, Total.Pupil.FTEs, Average.Expenditures.per.Pupil, X..Graduated, X..Still.in.School, X..Non.Grad.Completers, X..GED, X..Dropped.Out, X..Permanently.Excluded, X..Attending.College, MCAS_10thGrade_English_A.., MCAS_10thGrade_English_CPI, MCAS_10thGrade_Math_P.., MCAS_10thGrade_English_Incl..in.SGP..., MCAS_10thGrade_English_NI.., MCAS_10thGrade_English_P.., MCAS_10thGrade_English_P.A.., MCAS_10thGrade_English_SGP, MCAS_10thGrade_English_Stud..Incl..., MCAS_10thGrade_English_W.F.., MCAS_10thGrade_Math_A.., MCAS_10thGrade_Math_CPI, MCAS_10thGrade_Math_Incl..in.SGP..., MCAS_10thGrade_Math_NI.., MCAS_10thGrade_Math_P.., MCAS_10thGrade_Math_P.A.., MCAS_10thGrade_Math_SGP, MCAS_10thGrade_Math_Stud..Incl..., MCAS_10thGrade_Math_W.F.., X..MCAS_10thGrade_English_A, X..MCAS_10thGrade_English_NI, X..MCAS_10thGrade_English_P, X..MCAS_10thGrade_English_P.A, X..MCAS_10thGrade_English_W.F, X..MCAS_10thGrade_Math_A, X..MCAS_10thGrade_Math_NI, X..MCAS_10thGrade_Math_P, X..MCAS_10thGrade_Math_P.A, X..MCAS_10thGrade_Math_W.F)
```

Variables
Maggie: 1-22
Yue: 23 - 45
Emma: 46 - 67

##get rid of schools without tenth grade 
```{r}
subset(ma_schools_pruned,
       !is.na(ma_schools_pruned$MCAS_10thGrade_Math_A..))
```

```{r}
str(pruned_ma_schools)
```

#Key Next Steps
1. Eliminate the variables that we know off-hand are not useful in our analysis. 
2. Test for collinearity between variables that contain related information (for example, we will likely not include both Salary Totals and Average Salary, which both pertain to the salary of teachers).
3. Create a "Percent Proficient" variable to be used as our response variable out of the categories provided. 
4. Rename some of the variables so they make more sense.
